---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse) # A collection of packages - makes it easy to  load core packages in a single command.
library(RColorBrewer)#Color palettes
library(caTools)#collection of basic utility functions
library(party) #A computational toolbox for recursive partitioning.library(zoo) # for time series and ordered observations
library(janitor)##for creating frequency tables
library(caret) ## package for model training(Regression and classifications)
library(e1071) #Functions for latent class analysis
library(Boruta) #For relevant feature selection
library(tree) # Decision Tree Algorithm For Classification and Regression
library(rpart) #Recursive Partitioning and Tree Algorithm for classification and Regression
library(rattle) #GUI for better tree plots
library(rpart.plot)#Tree plotslibrary(lattice) #Graphics and visualizations
library(Metrics) #Evaluation metrics for Machine learning
library(kableExtra) #package for Creating complex tables
library(ggplot2)# package for creating charts/visualizations
library(ggpubr)# package for creating charts/visualizations
library(randomForest) #Random forest Machine Learning Algorithm
library(vtree)# Tree -like visualizations
library(AICcmodavg) #to calculate the average of the AIC
```
#Importing the datasets
```{r}
hot_1<- read_csv("H1.csv")
hot_2 <- read_csv("H2.csv")
```


#The data sets have the same number of columns and column names, they can therefore be joined along the row.
```{r}
#Using r-bind to join "hotel1" and "hotel2" and giving it a new variable name "hotel"

hotel_binded <- rbind(hot_1, hot_2)
```

#Confirming the data types of the variables based on documentation provided
```{r}
str(hotel_binded)
```

#Some variables are appearing as character when they should be categorical.
```{r}
#Converting the variables with wrong data types into the correct data type
hotel_binded$Agent <- as.factor(hotel_binded$Agent)
hotel_binded$IsCanceled <- as.factor(hotel_binded$IsCanceled)
hotel_binded$DistributionChannel <- as.factor(hotel_binded$DistributionChannel)  
hotel_binded$ReservedRoomType <- as.factor(hotel_binded$ReservedRoomType)
hotel_binded$DepositType <- as.factor(hotel_binded$DepositType)
hotel_binded$CustomerType <- as.factor(hotel_binded$CustomerType)
hotel_binded$Company <- as.factor(hotel_binded$Company)
hotel_binded$AssignedRoomType <- as.factor(hotel_binded$AssignedRoomType)
hotel_binded$ArrivalDateMonth <- as.factor(hotel_binded$ArrivalDateMonth)
hotel_binded$IsCanceled <- as.factor(hotel_binded$IsCanceled)
hotel_binded$IsRepeatedGuest <- as.factor(hotel_binded$IsRepeatedGuest)
hotel_binded$MarketSegment <- as.factor(hotel_binded$MarketSegment)
hotel_binded$Meal <- as.factor(hotel_binded$Meal)
hotel_binded$ReservationStatus <- as.factor(hotel_binded$ReservationStatus)
hotel_binded$ReservationStatusDate <- as.Date(hotel_binded$ReservationStatusDate)
hotel_binded$Country <- as.character(hotel_binded$Country)

#Checking to confirm that changes have been effected
str(hotel_binded)
```

#Checking for missing values
```{r}
sum(is.na(hotel_binded))
```

#There are 4 missing values in the data set in general. 

#Checking to confirm which columns have missing values
```{r}
names(which(colSums(is.na(hotel_binded)) > 0))
```

#Missing values are only found in the 'children' column.
```{r}
#Dealing with the missing values by replacing them with the median.

hotel_binded$Children[is.na(hotel_binded$Children)] <- median(hotel_binded$Children, na.rm=TRUE)

#Confirming that missing values have been dealt with
sum(is.na(hotel_binded))
```
######################DATA VISUALIZATION#######################
#What is the rate of Canceled and not Cancelled?
```{r}
xtabs(~IsCanceled, data=hotel_binded)
```


#More visualization of the rate of Canceled and not Canceled?
```{r}
vtree(hotel_binded, "IsCanceled")
```

#How long does it take for someone to cancel their bookings from when it was first booked?
```{r}
boxplot(LeadTime~IsCanceled,ylab="LeadTime", xlab="IsCanceled",col="red", data=hotel_binded)
```

#How many guests that booked are repeated guests?
```{r}
table(hotel_combined$IsRepeatedGuest)
```



#How many repeated guests canceled their reservations and how many maintained their reservations?
```{r}
xtabs(~IsCanceled+IsRepeatedGuest, data=hotel_binded)
```

#What year had the most cancellations?
```{r}
barplot(xtabs(~IsCanceled + ArrivalDateYear, data=hotel_binded), col=c ("red", "purple"), legend=c("Not Canceled", "Canceled"), beside=T)
```


#What Month had the most cancellations/most bookings?
```{r}
barplot(xtabs(~IsCanceled + ArrivalDateMonth, data=hotel_binded), col=c ("green", "blue"), legend=c(" Not Canceled", "Canceled"),las=2, beside=T)
```

#What market segment had the most impact on bookings?
```{r}
barplot(xtabs(~ReservationStatus+ MarketSegment, data=hotel_binded), col=c ("green","black","blue"), legend=c("Canceled", "CheckOut","No Show"),las=2, beside=T)
```


#### 1. Which country has the most number of hotel bookings/expected visitor?
```{r}
countrytabb= table(hotel_binded$Country) #create a frequency table for the country column
countrydff = data.frame(countrytabb) #create a dataframe for it
head(countrydff)
names(countrydff)[1] = 'Country' #Replacing the default column name in the dataframe with 'Country'
countrydff_new <- countrydff[order(countrydff$Freq),] #ordering the data frame by frequency in ascending order (default order)
countrydff_new
countriestopp <- tail(countrydff_new, n = 7) # creating a dataframe with the  7 countries that have the highest frequencies 
countriestopp
dim(countrydff_new) #To confirm total number of countries
sum(countriestopp$Freq) #to confirm the number of bookings from the top 7 countries
```
#What is the average daily rate per year?
```{r}
tapply(hotel_binded$ADR, hotel_binded$ArrivalDateYear, summary)
```

#This graph is plotted to see what deposit type would most likely cancel
```{r}
barplot(xtabs(~IsCanceled+DepositType, data=hotel_binded), col=c ("orange","red"), legend=c("Not Canceled", "Canceled"), beside=T)
```


```{r}
library(tidyverse)
  summary(hotel_binded, na.rm=T)
```


#To drop the Agent and Company Column, this is because of the NULL values in the dataset
```{r}
library(tidyverse)
bindednew<- subset(hotel_binded, select=-c(Agent,Company,ReservationStatusDate,ReservationStatus,ReservedRoomType,Meal,Adults,
                                           DistributionChannel,Children,PreviousBookingsNotCanceled,
                                           DaysInWaitingList,IsRepeatedGuest,Babies))
view(bindednew)
```

#Split the IsCanceled column into iss canceled and not canceled and assign them to a variable name
```{r}
h_data1 <- bindednew[which(bindednew$IsCanceled==0),]
h_data2 <- bindednew[which(bindednew$IsCanceled==1),]
```
#This ensures that we have equal amount of canceled and not canceled in  the training dataset
```{r}
set.seed(123)
train_part1 <- sample(1:nrow(h_data1),0.75*nrow(h_data1))
train_part2 <- sample(1:nrow(h_data2),0.75*nrow(h_data2))
#view(training_set1)
#view(training_set2)
```

#Put in a dataframe
```{r}
trainpart1 <- h_data1[train_part1,]
trainpart2 <- h_data2[train_part2,]
#trainpart1
#trainpart2
```

#Combine both dataframes
```{r}
trainbinded <- rbind(trainpart1,trainpart2)
#view(trainbinded)

```

#Get the train data set 
```{r}
testpart1 <- h_data1[-train_part1,] #contains 25% of those 0
testpart2 <- h_data2[-train_part2,]#contains 25% of those 1

testbinded <- rbind(testpart1, testpart2)
#view(testdata)
```

#Feature Importance
```{r}
library(randomForest)


forest_results <- randomForest(IsCanceled~., data=trainbinded)

print(forest_results)

#str(trainbinded)
```

#Feature selection/importance
```{r}
random_dfs <- randomForest::importance(forest_results)
random_dfs

randomf_dfs <- data.frame(Variables=row.names(random_dfs), MeanDecreaseGini=random_dfs[,1])
randomf_dfs <- random_dfs[order(randomf_dfs$MeanDecreaseGini, decreasing=TRUE),]
randomf_dfs
```

######################## RANDOM FOREST ALGORITHM ############################
```{r}
library(randomForest)
r_model1 <- randomForest(IsCanceled~.,data=trainbinded, importance=TRUE)
r_model1

#randomForest::importance(r_model1) 

```

```{r}
colnames(trainbinded)
```
```{r}
unique(trainbinded$ArrivalDateYear)
unique(trainbinded$ArrivalDateMonth)
unique(trainbinded$ArrivalDateWeekNumber)
unique(trainbinded$ArrivalDateDayOfMonth)
unique(trainbinded$StaysInWeekendNights)
unique(trainbinded$StaysInWeekNights)
unique(trainbinded$Country)
unique(trainbinded$MarketSegment)
unique(trainbinded$PreviousCancellations)
unique(trainbinded$AssignedRoomType)
unique(trainbinded$BookingChanges)
unique(trainbinded$DepositType)
unique(trainbinded$CustomerType)
unique(trainbinded$ADR)
unique(trainbinded$RequiredCarParkingSpaces)
unique(trainbinded$TotalOfSpecialRequests)
```
```{r}
str(trainbinded)
summary(trainbinded)
```


#####Accuracy
```{r}
accuracy_score <- (52484+26783)/(52484+3890+26783+6385) 
print(accuracy_score)
```
####Precision
```{r}
precision <- (26783)/(26783+3890)
print(precision)
```

####Recall
```{r}
recall <- (26783)/(26783+6385)
print(recall)
```
####F1 Score
```{r}
f1_Score <- (2*((0.87*0.80)/(0.87+0.80)))
print(f1_Score)
```

#Predict the train data
```{r}
library(tidyverse)
prediction.test <- predict(r_model1,testbinded,type="class")
view(prediction.test)
```

#Check classification accuracy
```{r}
testbinded$prediction.test <- prediction.test
view(testbinded)
```

```{r}
view(trainbinded)
```

```{r}
confusionMatrix(prediction.test, testbinded$IsCanceled) #Using a confusion matrix to examine the accuracy of the model.
```

```{r}
# Save model to RDS file
saveRDS(r_model1, "r_model1.rds")
```

